---
title: "Round 3 Voucher Market Report"
output:
  html_document:
    toc: true
    toc_float: true
    number_sections: true
    df_print: paged
params:
  bucket_size: 10000
  output_dir: "output"
---

```{r setup, include=FALSE}
user_library <- trimws(path.expand(Sys.getenv("R_LIBS_USER")))
if (nzchar(user_library) && dir.exists(user_library)) {
  .libPaths(unique(c(user_library, .libPaths())))
}

required_packages <- c(
  "dplyr",
  "ggplot2",
  "knitr",
  "purrr",
  "readr",
  "scales",
  "stringr",
  "tibble",
  "tidyr"
)

missing_packages <- required_packages[
  !vapply(required_packages, requireNamespace, logical(1), quietly = TRUE)
]

if (length(missing_packages) > 0) {
  stop(
    "Missing packages: ",
    paste(missing_packages, collapse = ", ")
  )
}

invisible(lapply(required_packages, library, character.only = TRUE))

knitr::opts_chunk$set(
  echo = FALSE,
  warning = FALSE,
  message = FALSE,
  fig.width = 14,
  fig.height = 6,
  dpi = 120
)

current_input <- knitr::current_input()
report_dir <- if (length(current_input) == 1 && nzchar(current_input)) {
  dirname(normalizePath(current_input, winslash = "/", mustWork = TRUE))
} else {
  normalizePath(getwd(), winslash = "/", mustWork = TRUE)
}

data_dir <- file.path(report_dir, "Data Capsule")
output_dir <- if (grepl("^(?:[A-Za-z]:|/|\\\\\\\\)", params$output_dir)) {
  params$output_dir
} else {
  file.path(report_dir, params$output_dir)
}
png_dir <- file.path(output_dir, "png")

dir.create(output_dir, recursive = TRUE, showWarnings = FALSE)
dir.create(png_dir, recursive = TRUE, showWarnings = FALSE)

bucket_size <- params$bucket_size

roll_mean <- function(x, window = 5) {
  if (length(x) == 0 || window <= 1) {
    return(x)
  }

  as.numeric(stats::filter(x, rep(1 / window, window), sides = 1))
}

nan_to_na <- function(x) {
  x[is.nan(x)] <- NA_real_
  x
}

read_capsule_file <- function(path, kind = c("prices", "trades")) {
  kind <- match.arg(kind)
  day_from_name <- as.integer(stringr::str_extract(basename(path), "(?<=day_)\\d+"))

  data <- readr::read_delim(
    path,
    delim = ";",
    show_col_types = FALSE,
    progress = FALSE,
    trim_ws = TRUE
  )

  if (!"day" %in% names(data)) {
    data <- dplyr::mutate(data, day = day_from_name)
  }

  if (kind == "trades" && !"product" %in% names(data)) {
    data <- dplyr::mutate(data, product = symbol)
  }

  dplyr::mutate(
    data,
    day = as.integer(day),
    global_time = day * 1000000 + timestamp
  )
}

price_files <- list.files(
  data_dir,
  pattern = "^prices_round_3_day_\\d+\\.csv$",
  full.names = TRUE
)
trade_files <- list.files(
  data_dir,
  pattern = "^trades_round_3_day_\\d+\\.csv$",
  full.names = TRUE
)

prices <- purrr::map_dfr(price_files, read_capsule_file, kind = "prices") %>%
  dplyr::arrange(product, global_time)

trades <- purrr::map_dfr(trade_files, read_capsule_file, kind = "trades") %>%
  dplyr::arrange(product, global_time)

day_breaks <- tibble::tibble(
  x = sort(unique(prices$day)) * 1000000
) %>%
  dplyr::filter(x > min(x))

day_labels <- tibble::tibble(
  day = sort(unique(prices$day)),
  start_x = sort(unique(prices$day)) * 1000000,
  mid_x = sort(unique(prices$day)) * 1000000 + 500000,
  label = paste0("day", sort(unique(prices$day)))
)

bucketed_prices <- prices %>%
  dplyr::mutate(
    bucket = floor(global_time / bucket_size),
    bucket_time = bucket * bucket_size + bucket_size / 2
  ) %>%
  dplyr::group_by(product, day, bucket, bucket_time) %>%
  dplyr::summarise(
    open = dplyr::first(mid_price),
    high = max(mid_price, na.rm = TRUE),
    low = min(mid_price, na.rm = TRUE),
    close = dplyr::last(mid_price),
    rel_spread = mean((ask_price_1 - bid_price_1) / pmax(mid_price, 1e-9), na.rm = TRUE),
    .groups = "drop"
  ) %>%
  dplyr::mutate(rel_spread = nan_to_na(rel_spread))

bucketed_trades <- trades %>%
  dplyr::mutate(
    bucket = floor(global_time / bucket_size),
    bucket_time = bucket * bucket_size + bucket_size / 2
  ) %>%
  dplyr::group_by(product, day, bucket, bucket_time) %>%
  dplyr::summarise(
    trade_count = dplyr::n(),
    trade_volume = sum(quantity, na.rm = TRUE),
    trade_notional = sum(price * quantity, na.rm = TRUE),
    .groups = "drop"
  )

spot <- bucketed_prices %>%
  dplyr::filter(product == "VELVETFRUIT_EXTRACT") %>%
  dplyr::arrange(bucket_time) %>%
  dplyr::mutate(
    rolling_close = roll_mean(close, 5)
  )

underlyings <- bucketed_prices %>%
  dplyr::filter(product %in% c("HYDROGEL_PACK", "VELVETFRUIT_EXTRACT")) %>%
  dplyr::arrange(product, bucket_time) %>%
  dplyr::group_by(product) %>%
  dplyr::mutate(
    rolling_close = roll_mean(close, 5)
  ) %>%
  dplyr::ungroup()

underlying_summary <- underlyings %>%
  dplyr::left_join(
    bucketed_trades %>%
      dplyr::filter(product %in% c("HYDROGEL_PACK", "VELVETFRUIT_EXTRACT")) %>%
      dplyr::group_by(product) %>%
      dplyr::summarise(
        total_trade_count = sum(trade_count, na.rm = TRUE),
        total_trade_volume = sum(trade_volume, na.rm = TRUE),
        total_trade_notional = sum(trade_notional, na.rm = TRUE),
        .groups = "drop"
      ),
    by = "product"
  ) %>%
  dplyr::group_by(product) %>%
  dplyr::summarise(
    start_price = dplyr::first(close),
    end_price = dplyr::last(close),
    net_change = dplyr::last(close) - dplyr::first(close),
    pct_change = 100 * (dplyr::last(close) / dplyr::first(close) - 1),
    min_price = min(close, na.rm = TRUE),
    max_price = max(close, na.rm = TRUE),
    price_range = max(close, na.rm = TRUE) - min(close, na.rm = TRUE),
    avg_rel_spread_pct = mean(rel_spread, na.rm = TRUE) * 100,
    total_trade_count = dplyr::first(total_trade_count),
    total_trade_volume = dplyr::first(total_trade_volume),
    total_trade_notional = dplyr::first(total_trade_notional),
    .groups = "drop"
  ) %>%
  dplyr::mutate(
    dplyr::across(
      c(
        start_price,
        end_price,
        net_change,
        pct_change,
        min_price,
        max_price,
        price_range,
        avg_rel_spread_pct
      ),
      ~ round(.x, 2)
    ),
    avg_trade_size = round(total_trade_volume / pmax(total_trade_count, 1), 2),
    avg_trade_price = round(total_trade_notional / pmax(total_trade_volume, 1), 2),
    dplyr::across(
      c(total_trade_volume, total_trade_notional),
      ~ round(.x, 0)
    )
  )

vouchers <- bucketed_prices %>%
  dplyr::filter(stringr::str_detect(product, "^VEV_")) %>%
  dplyr::mutate(
    strike = as.integer(stringr::str_extract(product, "\\d+"))
  ) %>%
  dplyr::left_join(
    dplyr::select(spot, bucket, spot_close = close),
    by = "bucket"
  ) %>%
  dplyr::left_join(
    dplyr::select(bucketed_trades, product, bucket, trade_count, trade_volume, trade_notional),
    by = c("product", "bucket")
  ) %>%
  dplyr::mutate(
    trade_count = tidyr::replace_na(trade_count, 0L),
    trade_volume = tidyr::replace_na(trade_volume, 0),
    trade_notional = tidyr::replace_na(trade_notional, 0),
    spot_minus_strike = spot_close - strike,
    intrinsic_value = pmax(spot_minus_strike, 0),
    time_value = close - intrinsic_value,
    rel_spread_pct = rel_spread * 100
  )

last_spot_close <- dplyr::last(spot$close)

voucher_summary <- vouchers %>%
  dplyr::group_by(product, strike) %>%
  dplyr::summarise(
    start_price = dplyr::first(close),
    end_price = dplyr::last(close),
    net_change = dplyr::last(close) - dplyr::first(close),
    pct_change = 100 * (dplyr::last(close) / dplyr::first(close) - 1),
    avg_rel_spread_pct = mean(rel_spread_pct, na.rm = TRUE),
    total_trade_count = sum(trade_count, na.rm = TRUE),
    total_trade_volume = sum(trade_volume, na.rm = TRUE),
    total_trade_notional = sum(trade_notional, na.rm = TRUE),
    avg_time_value = mean(time_value, na.rm = TRUE),
    last_time_value = dplyr::last(time_value),
    last_intrinsic_value = dplyr::last(intrinsic_value),
    .groups = "drop"
  ) %>%
  dplyr::mutate(
    dplyr::across(
      c(
        start_price,
        end_price,
        net_change,
        pct_change,
        avg_rel_spread_pct,
        avg_time_value,
        last_time_value,
        last_intrinsic_value
      ),
      nan_to_na
    ),
    last_spot = last_spot_close,
    distance_to_last_spot = strike - last_spot_close,
    avg_trade_size = total_trade_volume / pmax(total_trade_count, 1),
    avg_trade_price = total_trade_notional / pmax(total_trade_volume, 1)
  ) %>%
  dplyr::arrange(strike)

most_traded <- voucher_summary %>%
  dplyr::slice_max(order_by = total_trade_count, n = 1)

tightest_name <- voucher_summary %>%
  dplyr::filter(is.finite(avg_rel_spread_pct)) %>%
  dplyr::slice_min(order_by = avg_rel_spread_pct, n = 1)

richest_time_value <- voucher_summary %>%
  dplyr::slice_max(order_by = avg_time_value, n = 1)

base_theme <- ggplot2::theme_minimal(base_size = 12) +
  ggplot2::theme(
    plot.title = ggplot2::element_text(face = "bold", size = 15),
    plot.subtitle = ggplot2::element_text(size = 10, colour = "#475569"),
    panel.grid.minor = ggplot2::element_blank(),
    legend.position = "bottom"
  )

time_scale <- ggplot2::scale_x_continuous(
  labels = scales::label_number(scale_cut = scales::cut_short_scale())
)

add_day_lines <- function(plot) {
  plot +
    ggplot2::geom_vline(
      data = day_breaks,
      ggplot2::aes(xintercept = x),
      inherit.aes = FALSE,
      linetype = "dashed",
      colour = "#94a3b8",
      linewidth = 0.4
    )
}

add_day_labels <- function(plot) {
  plot +
    ggplot2::geom_text(
      data = day_labels,
      ggplot2::aes(x = mid_x, y = Inf, label = label),
      inherit.aes = FALSE,
      vjust = 1.4,
      colour = "#64748b",
      size = 3.4,
      fontface = "bold"
    )
}

save_plot <- function(plot, filename, width = 14, height = 6) {
  ggplot2::ggsave(
    filename = file.path(png_dir, filename),
    plot = plot,
    width = width,
    height = height,
    dpi = 320,
    bg = "white"
  )
}

underlyings_plot <- ggplot2::ggplot(underlyings, ggplot2::aes(x = bucket_time, y = close)) +
  ggplot2::geom_line(colour = "#93c5fd", linewidth = 0.35) +
  ggplot2::geom_line(
    ggplot2::aes(y = rolling_close),
    colour = "#2563eb",
    linewidth = 0.9
  ) +
  ggplot2::facet_wrap(~ product, ncol = 1, scales = "free_y") +
  ggplot2::labs(
    title = "Underlying price series",
    subtitle = paste0(
      "Bucket size: ",
      scales::comma(bucket_size),
      ". Thin lines show bucket closes. Thick lines show a 5-bucket rolling mean. Dashed lines and labels separate day0, day1, and day2."
    ),
    x = NULL,
    y = "Mid price"
  ) +
  time_scale +
  base_theme

underlyings_plot <- add_day_lines(underlyings_plot)
underlyings_plot <- add_day_labels(underlyings_plot)

strike_levels <- sort(unique(vouchers$strike))
spot_low <- min(spot$low, na.rm = TRUE)
spot_high <- max(spot$high, na.rm = TRUE)
visible_strikes <- strike_levels[
  strike_levels >= (spot_low - 35) &
    strike_levels <= (spot_high + 35)
]
if (length(visible_strikes) == 0) {
  visible_strikes <- strike_levels
}
y_min <- min(c(spot_low, visible_strikes), na.rm = TRUE) - 12
y_max <- max(c(spot_high, visible_strikes), na.rm = TRUE) + 12
right_edge <- max(spot$bucket_time) + bucket_size * 8

spot_strike_plot <- ggplot2::ggplot(spot, ggplot2::aes(x = bucket_time, y = close)) +
  ggplot2::geom_hline(
    data = tibble::tibble(strike = visible_strikes),
    ggplot2::aes(yintercept = strike),
    inherit.aes = FALSE,
    colour = "#94a3b8",
    linetype = "dotted",
    linewidth = 0.55
  ) +
  ggplot2::geom_line(colour = "#93c5fd", linewidth = 0.35) +
  ggplot2::geom_line(ggplot2::aes(y = rolling_close), colour = "#2563eb", linewidth = 0.95) +
  ggplot2::geom_text(
    data = tibble::tibble(strike = visible_strikes, label_x = right_edge),
    ggplot2::aes(x = label_x, y = strike, label = paste0("K=", strike)),
    inherit.aes = FALSE,
    hjust = 0,
    size = 3,
    colour = "#64748b"
  ) +
  ggplot2::labs(
    title = "VELVETFRUIT_EXTRACT versus voucher strike ladder",
    subtitle = "Horizontal lines mark the visible strike range around the spot path. Dashed lines separate day0, day1, and day2.",
    x = "Continuous time",
    y = "VELVETFRUIT_EXTRACT mid price"
  ) +
  ggplot2::coord_cartesian(ylim = c(y_min, y_max), clip = "off") +
  ggplot2::scale_x_continuous(
    labels = scales::label_number(scale_cut = scales::cut_short_scale()),
    expand = ggplot2::expansion(mult = c(0.01, 0.08))
  ) +
  base_theme

spot_strike_plot <- add_day_lines(spot_strike_plot)
spot_strike_plot <- add_day_labels(spot_strike_plot)

time_value_cap <- stats::quantile(vouchers$time_value, 0.95, na.rm = TRUE)
vouchers_for_plot <- vouchers %>%
  dplyr::mutate(time_value_plot = pmin(time_value, time_value_cap))

time_value_plot <- ggplot2::ggplot(
  vouchers_for_plot,
  ggplot2::aes(
    x = bucket_time,
    y = factor(strike, levels = rev(sort(unique(strike)))),
    fill = time_value_plot
  )
) +
  ggplot2::geom_tile() +
  ggplot2::scale_fill_gradientn(
    colours = c("#0f172a", "#4c1d95", "#db2777", "#f59e0b", "#fef08a"),
    na.value = "#f8fafc"
  ) +
  ggplot2::labs(
    title = "Voucher time value",
    subtitle = "Time value is calculated as voucher mid price minus intrinsic value. Colour scale is clipped at the 95th percentile to increase contrast.",
    x = "Continuous time",
    y = "Strike",
    fill = "Time value"
  ) +
  time_scale +
  base_theme

time_value_plot <- add_day_lines(time_value_plot)
time_value_plot <- add_day_labels(time_value_plot)

liquidity_plot <- ggplot2::ggplot(
  voucher_summary,
  ggplot2::aes(
    x = avg_rel_spread_pct,
    y = total_trade_count,
    size = total_trade_volume,
    colour = avg_time_value
  )
) +
  ggplot2::geom_point(alpha = 0.9) +
  ggplot2::geom_text(
    ggplot2::aes(label = product),
    hjust = -0.05,
    nudge_x = 2,
    size = 3,
    check_overlap = TRUE
  ) +
  ggplot2::scale_colour_viridis_c(option = "D", na.value = "#64748b") +
  ggplot2::scale_size_area(max_size = 14) +
  ggplot2::coord_cartesian(clip = "off") +
  ggplot2::expand_limits(x = max(voucher_summary$avg_rel_spread_pct, na.rm = TRUE) * 1.18) +
  ggplot2::labs(
    title = "Voucher liquidity summary",
    subtitle = "X-axis: average relative spread. Y-axis: total trade count. Bubble size: total traded volume. Colour: average time value.",
    x = "Average relative spread (%)",
    y = "Total trade count",
    size = "Trade volume",
    colour = "Avg time value"
  ) +
  base_theme

voucher_summary_table <- voucher_summary %>%
  dplyr::mutate(
    start_price = round(start_price, 2),
    end_price = round(end_price, 2),
    net_change = round(net_change, 2),
    pct_change = round(pct_change, 2),
    avg_rel_spread_pct = round(avg_rel_spread_pct, 2),
    total_trade_volume = round(total_trade_volume, 0),
    total_trade_notional = round(total_trade_notional, 0),
    avg_trade_size = round(avg_trade_size, 2),
    avg_trade_price = round(avg_trade_price, 2),
    avg_time_value = round(avg_time_value, 2),
    last_time_value = round(last_time_value, 2),
    last_intrinsic_value = round(last_intrinsic_value, 2),
    last_spot = round(last_spot, 2),
    distance_to_last_spot = round(distance_to_last_spot, 2),
    share_of_trade_count_pct = round(
      100 * total_trade_count / sum(total_trade_count, na.rm = TRUE),
      2
    )
  )

save_plot(underlyings_plot, "01_underlying_price_series.png", width = 14, height = 8)
save_plot(spot_strike_plot, "02_spot_vs_strikes.png")
save_plot(time_value_plot, "03_time_value_heatmap.png")
save_plot(liquidity_plot, "04_liquidity_map.png", width = 14, height = 7)
```

# Report Scope

This report summarizes Round 3 price, strike, time value, and liquidity structure using bucketed data.

Bucket size: `r scales::comma(bucket_size)` time steps.

# Summary Statistics

- Most traded voucher: **`r most_traded$product`** (`r scales::comma(most_traded$total_trade_count)` trades)
- Lowest average relative spread: **`r tightest_name$product`** (`r round(tightest_name$avg_rel_spread_pct, 2)`%)
- Highest average time value: **`r richest_time_value$product`** (`r round(richest_time_value$avg_time_value, 2)`)

# 1. Underlying Price Series

The chart below shows bucketed price series for `HYDROGEL_PACK` and `VELVETFRUIT_EXTRACT`. Thin lines show bucket closes. Thick lines show a 5-bucket rolling mean.

```{r underlyings}
underlyings_plot
```

# Underlying Summary Table

```{r underlying-summary-table}
knitr::kable(
  underlying_summary,
  caption = "Underlying summary for HYDROGEL_PACK and VELVETFRUIT_EXTRACT."
)
```

# 2. VELVETFRUIT_EXTRACT Versus Strike Ladder

The chart below shows `VELVETFRUIT_EXTRACT` against the voucher strike ladder. Horizontal reference lines mark strike levels.

```{r spot-vs-strikes}
spot_strike_plot
```

# 3. Voucher Time Value

The heatmap below shows voucher time value by strike and time bucket. Time value is defined as voucher mid price minus intrinsic value.

```{r time-value-heatmap}
time_value_plot
```

# 4. Voucher Liquidity Summary

The chart below summarizes voucher liquidity. Lower values on the x-axis indicate tighter average spreads. Higher values on the y-axis indicate more trade activity.

```{r liquidity-map}
liquidity_plot
```

# Voucher Summary Table

```{r summary-table}
knitr::kable(
  voucher_summary_table,
  caption = "Round 3 voucher summary."
)
```
